---
title: "CodingChallenege6"
author: "Vaibhav B. Shelar"
date: "2025-03-27"
output: 
      html_document: 
      variant: gfm
      md_document:
      word_document: default
      toc: true
      toc_float: true
      pdf_document: default
---
### Q1

# Regarding reproducibility, what is the main point of writing your own functions and iterations? 

# Ans: The main point of writing your own functions and iterations is to make the code more readable, reusable, and reproducible.

### Q2

# In your own words, describe how to write a function and a for loop in R and how they work. Give me specifics like syntax, where to write code, and how the results are returned. 

# Ans: To write a function in R, we use the `function` keyword followed by the function name and its arguments. The body of the function is enclosed in curly braces `{}` and contains the code that performs the desired task.

# Examples

# 1)
```{r}
C_to_F <- function(c_temp) {
  fahrenheit <- (c_temp * 9/5) + 32  # Conversion formula
  return(fahrenheit)  # Return the result
}

# Example usage:
C_to_F(25)  
```


```{r}
for(i in -25:50){
  result <- C_to_F(i)
  print(result)
}
```


### Q3 
#Read in the Cities.csv file from Canvas using a relative file path.

```{r}
# Read in the dataset
cities <- read.csv("Cities.csv")
```

### Q4

```{r}
# Function to calculate the distance between two pairs of coordinates
haversine_distance <- function(lat1, lon1, lat2, lon2) {
  rad.lat1 <- lat1 * pi/180  # convert to radians
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  delta_lat <- rad.lat2 - rad.lat1 # Haversine formula
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a)) 
  earth_radius <- 6378137 # Earth's radius in kilometers
  distance_km <- (earth_radius * c)/1000 # Calculate the distance
  return(distance_km)
}
```

### Q5

# Using your function, compute the distance between Auburn, AL and New York City 
# a.	Subset/filter the Cities.csv data to include only the latitude and longitude values you need and input as input to your function. 

# a

```{r}
library(dplyr)
# Subset the data for Auburn, AL and New York City
nyc<- cities %>% filter(city == "New York") 
auburn<- cities %>% filter(city == "Auburn")

# Extracting just latitude and longitude values
lat1<- nyc$lat
lon1<- nyc$long

lat2<- auburn$lat
lon2<- auburn$long

# b 

haversine_distance(lat1, lon1, lat2, lon2)

```

### Q6

# Now, use your function within a for loop to calculate the distance between all other cities in the data

```{r}
# loops to calculate distance from auburn to each city
for(i in 1:nrow(cities)){
  if(cities$city[[1]] != "Auburn"){
    lat1<- cities$lat[i]
    lon1<- cities$long[i]
    
    dist<- haversine_distance(lat1, lon1, lat2, lon2)
    print(dist)
  }
}

```
## Bonus point code

```{r}
# creating a empty data frame to store results
citydistance_df<- data.frame(city1 = character(),
                         city2 = character(),
                         distance_km = numeric())
# Loop through each city
for(i in 1:nrow(cities)){
  if(cities$city[i] != "Auburn"){
    lat1<- cities$lat[i]
    lon1<- cities$long[i]
# calculating the distance
    dist<- haversine_distance(lat1, lon1, lat2, lon2)
    
# Adding results as new row
    new_row<- data.frame(city1 = cities$city[i],
                         city2 = "Auburn",
                         distance_km = dist)
    
    citydistance_df<- rbind(citydistance_df, new_row)
  }
  
}

# first six rows of the data frame
head(citydistance_df)
```


### Q7 Adding a link to my github

(Github_Link)[https://github.com/vbs0303/Class-Reports.git]




