---
title: "R Markdown Coding Challenge"
author: "Vaibhav Shelar"
date: "2025-02-27"
output:
  md_document:
    variant: gfm
  pdf_document: default
  html_document: 
    toc: true
    toc_float: true
  word_document: default
---

- [Link of manuscript](https://doi.org/10.1094/PDIS-06-21-1253-RE)

- [GitHub flavored .md file](R-Markdown-Coding-Challenge)

## File Tree

```{r}
library("fs")
fs::dir_tree(path = "D:/Spring 2025/PLPA7820/Class-Reports/Assignments/Coding Challnge 4")
```


## Q1
# YAML Header: 
YAML header is a short blob of text, specially formatted with key: 
value pairs tags, that seats at the top of our Rmarkdown document. It contains 
YAML arguments such as "Title", "Author", and "Output", demarcated by three 
dashes (---) on either end. 

# Literate Programming: 
A methodology that combines a programming language with
a documentation language, thereby making programs more robust, more portable, 
more easily maintained, and arguably more fun to write than programs that are 
written only in a high-level language.


## Q2

```{r}
# Load dataset of MycotoxinData
Mycotoxin_Data <- read.csv("D:/Spring 2025/PLPA7820/Class-Reports/MycotoxinData.csv", na.strings = "na")
```

```{r}
# structure of the dataframe
str(Mycotoxin_Data)
```

```{r}
# plotting data using ggplota2 (Boxplot)

library(ggplot2)

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
library(ggplot2)
library(ggpubr)
```


```{r}
DONplot.pwc <- ggplot(Mycotoxin_Data, aes(x = Treatment, y = DON, fill = Cultivar)) +
  geom_boxplot(outliers = F) + 
  xlab("") + 
  ylab("DON (ppm)") + 
  geom_point(alpha = 0.6, pch = 21, color = "black", position = position_jitterdodge()) +
  scale_color_manual(values = c(cbbPalette[3], cbbPalette[4])) +
  scale_fill_manual(values = c(cbbPalette[3], cbbPalette[4])) +
  theme_classic() +
  facet_wrap(~Cultivar) +
  stat_compare_means(method = "t.test", label = "p.signif", 
                   comparisons = list(c("NTC", "Fg"), 
                                      c("Fg", "Fg + 37"), 
                                      c("Fg + 37", "Fg + 40"), 
                                      c("Fg + 40", "Fg + 70")))

DONplot.pwc
```

```{r}
X15ADONplot.pwc <- ggplot(Mycotoxin_Data, aes(x = Treatment, y = X15ADON, fill = Cultivar)) +
  geom_boxplot(outlier.shape = NA) + 
  xlab("") + 
  ylab("15ADON") + 
  geom_point(alpha = 0.6, pch = 21, color = "black", position = position_jitterdodge()) +
  scale_color_manual(values = c(cbbPalette[3], cbbPalette[4])) +
  scale_fill_manual(values = c(cbbPalette[3], cbbPalette[4])) +
  theme_classic() +
  facet_wrap(~Cultivar) +
  stat_compare_means(method = "t.test", label = "p.signif", 
                     comparisons = list(
                       c("NTC", "Fg"), 
                       c("Fg", "Fg + 37"), 
                       c("Fg + 37", "Fg + 40"), 
                       c("Fg + 40", "Fg + 70")
                     ))  

X15ADONplot.pwc
```


```{r}
Seed.massplot.pwc <- ggplot(Mycotoxin_Data, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar)) +
  geom_boxplot(outliers = F) + 
  xlab("") + 
  ylab("Seed Mass (mg)") + 
  geom_point(alpha = 0.6, pch = 21, color = "black", position = position_jitterdodge()) +
  scale_color_manual(values = c(cbbPalette[3], cbbPalette[4])) +
  scale_fill_manual(values = c(cbbPalette[3], cbbPalette[4])) +
  theme_classic() +
  facet_wrap(~Cultivar) +
  stat_compare_means(method = "t.test", label = "p.signif", 
                     comparisons = list(c("NTC", "Fg"), 
                                        c("Fg", "Fg + 37"), 
                                        c("Fg + 37", "Fg + 40"), 
                                        c("Fg + 40", "Fg + 70")))

Seed.massplot.pwc
```

```{r}
Combineplot.pwc <- ggarrange(
  DONplot.pwc,
  X15ADONplot.pwc, 
  Seed.massplot.pwc, 
  labels = c("A", "B", "C"),
  nrow = 1,
  ncol = 3,
  common.legend = TRUE,
  legend = "bottom"
)

Combineplot.pwc
```

