---
title: "Homework_DataWrangling"
author: "Vaibhav B. Shelar"
date: "2025-03-19"
output: 
  html_document: 
    keep_md: true
editor_options: 
  chunk_output_type: inline
---

## Seemless data wrangling

The tidyverse is a bunch of packages and functions written by the folks that manages Rstudio.
The tidyverse builds upon base R to allow for easier use of large datasets.h

## Install package tidyverse

install.packages("tidyverse")

```{r}
library(tidyverse)
```

```{r}
microbiome.fungi <- read.csv("Bull_richness.csv")
str(microbiome.fungi)
```

## select()

```{r}
microbiome.fungi2 <- select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
```

## filter()

```{r}
head(filter(microbiome.fungi2, Treatment == "Conv."))

# A more complex using & 
head(filter(microbiome.fungi2, Treatment == "Conv." & Fungicide == "C"))

# Another more complex using or |
head(filter(microbiome.fungi2, Sample == "A" | Sample == "B")) #Sample A or B

```
## mutate()

It allows us to quickly create new columns

```{r}
microbiome.fungi2$logRich <- log(microbiome.fungi2$richness)

# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich = log(richness)))

# Creating a new column which combines Crop and Treatment
head(mutate(microbiome.fungi2, Crop_Treatment = paste(Crop, Treatment)))
```

### the pipe '%>%'

```{r}
library(dplyr)
```


```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%  # selecting columns
  filter(Treatment == "Conv.") %>%  # subsetting to only include the conventional treatments
  mutate(logRich = log(richness)) %>%  # creating new column of the log richness
  head()  # displaying the forst six rows
```


### 'summerise()'

Use to find the mean, standard deviations/errors

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%  # selecting columns
  filter(Treatment == "Conv.") %>%  # subsetting to only include the conventional treatments
  mutate(logRich = log(richness)) %>%  # creating new column of the log richness
  summarise(Mean.rich = mean(logRich))  #Calculating overall mean log richness within the conventionally managed treatments 
```

### 'group_by()'

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%  # selecting columns
  filter(Treatment == "Conv.") %>%  # subsetting to only include the conventional treatments
  mutate(logRich = log(richness)) %>%  # creating new column of the log richness
  summarise(Mean.rich = mean(logRich),  #Calculating the mean richness, stdeviation, adn standard error
            n = n(),
            sd.dev = sd(logRich)) %>% 
  mutate(std.err = sd.dev/sqrt(n))
```

### Connecting to plotting

```{r}
library(ggplot2)
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%  # selecting columns
  group_by(Treatment, Fungicide) %>%  # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness 
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error 
            n = n(),
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>% 
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot 
  geom_bar(stat = "identity") +
  geom_errorbar(aes(x = Fungicide, ymin = Mean.rich - std.err, ymax = Mean.rich + std.err), width = 0.4) + 
  theme_minimal() + 
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```

### Joining

```{r}
# selecting just the richness and sample id 
richness <- microbiome.fungi %>% 
select(SampleID, richness)  

# selecting the columns that dont include the richness
metadata <- microbiome.fungi %>% 
select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(metadata)
head(richness)

head(left_join(metadata, richness, by = "SampleID")) # adding the richness data to the metadata based on the common column of sampleID
```

### Pivoting
Pivoting is also useful for converting from wide to long format and back again. We can do this with 
'pivot_longer()' and 'pivot_wider()'


```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%
   group_by(Treatment, Fungicide) %>%  # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness))  # Calculate the mean per Treatment and fungicide
```
# Wide format -dets the values within the fungicide column into column names

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%
   group_by(Treatment, Fungicide) %>%  # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>%  # Calculate the mean per Treatment and fungicide
  pivot_wider(names_from = Fungicide, values_from = Mean) # pivot to wide format
```


```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>%
   group_by(Treatment, Fungicide) %>%  # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>%  # Calculate the mean per Treatment and fungicide
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) # calculate the difference between means
```

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness")
```















